---
title: "National Attachment and Preferences on Redistribution"
author: "Noah Daponte-Smith"
date: "3/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(haven)
library(estimatr)
library(stargazer)
library(labelled)
library(modelsummary)
library(parlitools)
#read in data 

#bes <- read_dta("bes/internet_panel/bes_wave19_2019post.dta")
bes <- read_dta("bes_wave19_2019post.dta")
lookup <- read_csv("pcon_region_lookup_2019.csv")
#lookup <- read_csv("uk_geography/pcon_data/pcon_region_lookup_2019.csv")
#read in pcon to region lookup
```

```{r cleaning, echo = FALSE}
dat <- bes
dat[dat < 0] <- NA
dat[dat == 9999] <- NA
dat <- dat %>% 
  mutate(lad_name = labelled::to_factor(dat$oslaua),
         constit = labelled::to_factor(dat$pcon),
         PCON19NM = constit) %>% 
  left_join(., lookup, by = "PCON19NM") %>% 
  dplyr::select(id, wt, lad_name, constit, country, oslaua, gor, pcon, starts_with("PCON"), region, 
                jobzone, generalElectionVote, partyId, starts_with("p_"),
                starts_with("satDem"), leftRight, howLongLivedHere,
                age, gender, redistSelf, workingStatus, econPersonalProsp,
                starts_with("ukAttach"),
                starts_with("subnat")) %>% 
  mutate(vote_character = labelled::to_factor(dat$generalElectionVote)) %>% 
  filter(!(generalElectionVote %in% c(0, 9, 13, 9999))) %>% 
  mutate(vote_con = case_when(
    vote_character %in% c("Conservative", "Brexit Party","United Kingdom Independence Party (UKIP)") ~ 1,
    TRUE ~ 0
  ), vote = case_when(
    vote_character %in% c("Brexit Party","United Kingdom Independence Party (UKIP)",
                          "British National Party (BNP)") ~ 1, 
    vote_character %in% c("Conservative") ~ 2, 
    vote_character %in% c("Liberal Democrat", "Change UK- The Independent Group") ~ 3, 
    vote_character %in% c("Labour") ~ 4, 
    vote_character %in% c("Green Party","Plaid Cymru") ~ 5
  ), country = case_when(
    country == 1 ~ "England",
    country == 2 ~ "Scotland",
    country == 3 ~ "Wales"
  )) 

dat$age <- ifelse(dat$age < 0 | dat$age > 100, NA, dat$age)
dat$leftRight <- ifelse(dat$leftRight > 10, NA, dat$leftRight)
dat$p_gross_household <- ifelse(dat$p_gross_household > 15, NA, dat$p_gross_household)
dat$p_gross_personal <- ifelse(dat$p_gross_personal > 14, NA, dat$p_gross_personal)
dat$p_education <- ifelse(dat$p_education > 18, NA, dat$p_education)
dat$econPersonalProsp <- ifelse(dat$econPersonalProsp > 5, NA, dat$econPersonalProsp)
dat$redistSelf <- ifelse(dat$redistSelf > 10, NA, dat$redistSelf)
dat$p_ethnicity <- ifelse(dat$p_ethnicity > 14, NA, dat$p_ethnicity)
dat <- dat %>% filter(!is.na(p_ethnicity))
dat$is_white <- ifelse(dat$p_ethnicity == 1, 1, 0 )
dat$has_degree <- ifelse(dat$p_edlevel > 3, 1, 0)
dat$p_socgrade <- ifelse(dat$p_socgrade %in% c(7, 8), NA, dat$p_socgrade)

dat <- dat %>% 
  mutate(gross_household_mid = case_when(
    p_gross_household == 1 ~ 2500, 
    p_gross_household == 2 ~ 7500, 
    p_gross_household == 3 ~ 12500,
    p_gross_household == 4 ~ 17500,
    p_gross_household == 5 ~ 22500,
    p_gross_household == 6 ~ 27500,
    p_gross_household == 7 ~ 32500,
    p_gross_household == 8 ~ 37500,
    p_gross_household == 9 ~ 42500,
    p_gross_household == 10 ~ 47500,
    p_gross_household == 11 ~ 55000,
    p_gross_household == 12 ~ 65000,
    p_gross_household == 13 ~ 85000,
    p_gross_household == 14 ~ 125000,
    p_gross_household == 15 ~ 175000,
    TRUE ~ p_gross_household
  )) %>% 
  mutate(gross_household_mid = ifelse(gross_household_mid %in% c(16, 17), NA, gross_household_mid))

#get rid of p_ prefixes
names <- str_remove(names(dat), "p_")
names(dat) <- names

# reverse scale of attachment variables 4-6, to match directionality of 1-3
dat <- dat %>% 
  # reverse scale of 4-6 variables
  mutate(ukAttach4 = max(ukAttach4, na.rm = TRUE) - ukAttach4,
         ukAttach5 = max(ukAttach5, na.rm = TRUE) - ukAttach5,
         ukAttach6 = max(ukAttach6, na.rm = TRUE) - ukAttach6)

# do factor analysis of the 
attach <- dat %>% 
  select(id, starts_with("ukAttach")) %>% 
  na.omit()

# do the pca analysis 
fit <- princomp(attach[,2:7], cor=TRUE)
# plot pca - looks like they map to one component 
plot(fit, type = "lines")
# extract the first component 
# multiply by -1 so higher scores = more attached to UK
attach$uk_attach_score <- fit$scores[,1]*-1
# join first component back with the data
dat <- left_join(dat, attach %>% 
                   select(id, uk_attach_score), by = "id")

# get rid of variables about party contact - i don't care about them 
dat <- dat %>% 
  select(-starts_with("pContact"))

```

```{r}

# do factor analysiis osn subnational attachment variables 
dat <- dat %>% 
  mutate(subnatAttach4 = max(subnatAttach4, na.rm = TRUE),
         subnatAttach5 = max(subnatAttach5, na.rm = TRUE),
         subnatAttach6 = max(subnatAttach6, na.rm = TRUE))
subnat <- dat %>% 
  select(id, starts_with("subnatAttach")) %>% 
  na.omit()

subnat_fit <- princomp(subnat[,2:7], cor = TRUE)
plot(subnat_fit, type = "lines") # matches to one dimension

# extract first component 
# multiply by -1 so higher scores are more attached to subnational
subnat$subnat_attach_score <- subnat_fit$scores[,1]*-1
dat <- left_join(dat, 
                 subnat %>% select(id, subnat_attach_score), 
                 by = "id")

```
```{r context_variables}
# this chunk adds variables relating to local context 

# add variables on 2019 vote choice in constituency 
vote_vars <- bes_2019 %>% 
  select(ons_const_id:constituency_name,
         winner_19, con_19:other_19)
dat <- left_join(dat, vote_vars, by = c("constit" = "constituency_name",
                                         "PCON19CD" = "ons_const_id"))


# make long version of parties, that contains variable for each party's share

parties_long <- vote_vars %>% 
  pivot_longer(con_19:other_19,
               names_to = "party",
               values_to = "share") %>% 
  select(-winner_19) %>% 
  mutate(party = str_remove(party, "_19"),
         party = case_when(
           party == "con" ~ "Conservative",
           party == "lab" ~ "Labour",
           party == "ld" ~ "Liberal Democrat",
           party == "brexit" ~ "Brexit Party",
           party == "pc" ~ "Plaid Cymru",
           party == "snp" ~ "Scottish National Party (SNP)",
           party == "green" ~ "Green Party",
           TRUE ~ party
         )) %>% 
  rename(inparty_share = share) %>% 
  # scale the inparty_share vriable
  mutate(inparty_share_scale = scale(inparty_share))

# join this to dat by constituency and general election vote
# dat now has a variable for the constituency-level share of the vote for each voter's in-party
test <- left_join(dat, parties_long, 
                  by = c("constit" = "constituency_name",
                         "PCON19CD" = 'ons_const_id',
                         "ons_region_id",
                         "vote_character" = "party")) 


summary(lm(data = test %>% 
             filter(vote_character == "Conservative",
                    country == "England"), 
           redistSelf ~ edlevel + age + is_white + 
             socgrade + gross_household + inparty_share_scale))
```


## Redistribution attitudes and national attachment. 

I am interested in the relationship between individuals' group identities and their attitudes towards redistribution. As a preliminary analysis, I begin by considering senses of attachment to the nation. 

To create a measure of an individual's attachment to the nation, I use six survey items in the post-election 2019 wave of the British Election Study that ask respondents about the extent to which they agree with various statements about Britishness. I run a factor analysis on these six variables to extract the latent dimension onto which they map, and I use this extracted factor as my primary measure of national attachment.

My dependent variable is a survey item that asks respondents about their views on whether the government ought to try to equalize incomes. Smaller numbers indicate the government should do more to equalize incomes; positive coefficients on the variables indicate that the variable correlates with preferences for lower redistribution. 

```{r uk_attach}
# run a few models regressing redistribution on attachment

#model one - basic demographics
mod1 <- lm_robust(data = dat, 
           redistSelf ~ uk_attach_score + is_white + socgrade + 
             edlevel + age + gross_household)

# model 2 - add leftRight and party fixed effects
mod2 <- lm_robust(data = dat, redistSelf ~ uk_attach_score + is_white + socgrade + 
             edlevel + age + gross_household + leftRight, 
             fixed_effects = partyId * country)

modelsummary(list(mod1, mod2),
             stars = TRUE,
             coef_rename = c("uk_attach_score"= "UK Attachment", 
                             "is_white" = "White",
                             "socgrade" = "Social grade",
                             "edlevel" = "Education",
                             "age" = "Age",
                             "gross_household" = "income",
                             "leftRight" = "Left-Right"),
             coef_omit = "(Intercept)",
             notes = "Model 2 includes party fixed effects",
             title = "Redistribution preferences")

```

## Subnational attachment 

I look next at attachment to the subnational unit - in this case, the country (England, Scotland, and Wales). The BES' questions on subnational attachment are identical to those on national attachment; as such, I use the same factor-analysis strategy as above. While 
```{r subnat_attach}

mod_full <- lm_robust(data = dat, redistSelf ~ subnat_attach_score + is_white + socgrade + 
             edlevel + age + gross_household+ leftRight, 
             fixed_effects = ~country)

mod_england <- lm_robust(data = dat %>% filter(country == "England"),
                  redistSelf ~ subnat_attach_score + is_white + socgrade + 
                    edlevel + age + gross_household + leftRight, 
                  fixed_effects = partyId)

mod_celtic <- lm_robust(data = dat %>% filter(country %in% c( "Scotland", "Wales")),
                  redistSelf ~ subnat_attach_score + is_white + socgrade + 
                    edlevel + age + gross_household+ leftRight,
                  fixed_effects = partyId)

mod_scotland <- lm_robust(data = dat %>% filter(country %in% c( "Scotland")),
                  redistSelf ~ subnat_attach_score + is_white + socgrade + 
                    edlevel + age + gross_household + leftRight)

mods <- list(mod_full, 
             mod_england, mod_celtic, mod_scotland)

modelsummary(mods, 
             stars = TRUE)
```


