---
title: "Redistribution and Place-Based Economies"
author: "Noah Daponte-Smith"
date: "4/6/2022"
header-includes: 
  - \usepackage{setspace}\doublespacing
output: pdf_document

---

```{r setup, include=FALSE}
library(sf)
library(tidyverse)
library(stargazer)
library(estimatr)
library(modelsummary)
library(ggeffects)
library(patchwork)
library(parlitools)
library(DescTools)

knitr::opts_chunk$set(echo = TRUE)
 setwd("C:/Users/dapon/Dropbox/Harvard/dissertation/data")
 df <- read_csv("bes/internet_panel/clean_data/bes_wave21_clean.csv") 
 df$areaRichPoor_scale <- scale(df$areaRichPoor)[,1]
#df <- read_csv("bes_wave21_clean.csv")
# #read in region shapefile, rename regions
# shp <- st_read("uk_geography/shapefiles/NUTS_Level_1_(January_2018)_Boundaries.shp")
# # #change coordinate system to WGS 84
# shp <- st_transform(shp, "+proj=longlat +datum=WGS84") %>%
#   mutate(region = case_when(
#     nuts118cd == "UKC" ~ "North East",
#     nuts118cd == "UKD" ~ "North West",
#     nuts118cd == "UKE" ~ "Yorkshire and the Humber",
#     nuts118cd == "UKF" ~ "East Midlands",
#     nuts118cd == "UKG" ~ "West Midlands",
#     nuts118cd == "UKH" ~ "East of England",
#     nuts118cd == "UKI" ~ "London",
#     nuts118cd == "UKJ" ~ "South East",
#     nuts118cd == "UKK" ~"South West",
#     nuts118cd == "UKL" ~ "Wales",
#     nuts118cd == "UKM" ~ "Scotland",
#     nuts118cd == "UKN" ~ "Northern Ireland",
#   )) %>%
#   filter(region != "Northern Ireland")
```



## Regional deservingness and place-based economic resentment 
I first look at the relationship between preferences over redistribution (my dependent variable) and perceptions of the geographic distribution of central-government funding in Britain. The key independent variables are a series of questions that ask respondents about whether whether various geographic areas receive their "fair share" of central government spending. The scale is from 1-5, from "much less than its fair share" to "much more than its fair share." Respondents were asked about three geographic areas: their local area, their region, and London. I plot the distributions of these variables below. Most respondents believe their local areas and regions are under-funded relative to what they deserve, while the large majority believe London is overfunded. 

```{r histograms, echo = FALSE, warning = FALSE}
labels_custom <- c("much less", "a little less","about right","a little more", "much more")

local_hist <- df %>% 
  # have to do this factor thing to make the custom labels work correctly
  mutate(localFairShare_factor = factor(localFairShare)) %>% 
  filter(!is.na(localFairShare_factor)) %>% 
  ggplot() + 
  geom_histogram(aes(x = localFairShare_factor) , stat = "count") + 
  theme_minimal() + 
  labs(x = "Local area gets fair share of spending") +
  scale_x_discrete(labels = NULL)
  #scale_x_discrete(labels = labels_custom)

region_hist <- df %>% 
  mutate(regionFairShare_factor = factor(regionFairShare)) %>% 
  filter(!is.na(regionFairShare_factor)) %>% 
  ggplot() + 
  geom_histogram(aes(x = regionFairShare_factor), stat = "count") + 
  theme_minimal() + 
  labs(x = "Region gets fair share of spending",
       y = "") + 
  scale_x_discrete(labels = NULL)
  #scale_x_discrete(labels= labels_custom)

london_hist <- df %>% 
  mutate(londonFairShare_factor = factor(londonFairShare)) %>% 
  filter(!is.na(londonFairShare_factor)) %>% 
  ggplot() + 
  geom_histogram(aes(x = londonFairShare_factor), stat = "count") + 
  theme_minimal() + 
  labs(x = "London gets fair share of spending") + 
  scale_x_discrete(labels= labels_custom)

(local_hist | region_hist) / london_hist

```

I also plot the geographic distribution of these attitudes. The left and right panels plot the mean value by region of the regional-fair share and London-fair shares variables respectively. Scotland, Wales, London, and South East England are the only regions that rank above the median value (by region) of the regional fair-share variable. This accords with the facts that the latter two regions are either London or its immediate environs, while centralized redistribution to Scotland and Wales' devolved authorities has high political salience in British politics. The right panel shows that \textit{every} region, including London itself, believes London receives at least its proper share of central government spending, but that this feeling is most pronounced in northern England. Again, this accords with the prominent trend of north-south divides in British politics. 

```{r echo = FALSE, warning = FALSE}
region_fairshare <- df %>%
  group_by(region) %>%
  summarize(fairshare = mean(regionFairShare, na.rm = TRUE),
            n = n()) 

region_plot <- left_join(shp, region_fairshare,
           by = "region") %>%
  rename(`Region fair share` = fairshare) %>% 
  ggplot() +
  geom_sf(aes(fill = `Region fair share`)) +
  theme(rect = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(  ),
        text = element_text(size = 10),
        legend.position = "left",
        plot.title = element_text(hjust = 0),
        plot.margin = margin(0.1, 0, 0.1, 0, "cm")
  )  +
  scale_fill_gradient2(low = ("blue"), mid = "white",
                     high = ("red"), midpoint = median(region_fairshare$fairshare))

london_fairshare <- df %>% 
  group_by(region) %>% 
  summarize(`London fair share` = mean(londonFairShare, na.rm = TRUE)) 

london_plot <- left_join(shp, london_fairshare, 
                         by = "region") %>% 
  ggplot() + 
  geom_sf(aes(fill = `London fair share`)) + 
  theme(rect = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        text = element_text(size = 10),
        legend.position = "right", 
        plot.title = element_text(hjust = 0),
        plot.margin = margin(0.1, 0, 0.1, 0, "cm")) + 
  scale_fill_gradient2(low = ("blue"), mid = "white",
                       high = ("red"), midpoint = median(london_fairshare$`London fair share`))


region_plot + london_plot
```


## Relationship with preferences over redistribution
I now look at the relationship between these variables and a political outcome - preferences over redistriubtion. In eahc model, the key dependent variable is a 1-10 scale; higher values indicate preferences for higher levels of redistribution. Models 1 and 2 use the local and regional measures respectively as independent variables. For models 3 and 4, the dependent variable is the difference between the respondent's response to the London question and the local area and region questions respectively. As such, these latter models assess the respondent's perception that London's share of central government spending exceeds that of their own local area or region. This might plausibly capture a sense of place-based distributional resentment. I use linear models in each case. I include regional fixed effects, alongside a suite of demographic covariates - age, gender, social grade, education, household income, and the respondent's left-right placement. Positive coefficients indicate higher values of the variable correspond with preferences for higher redistribution. 
\newpage


```{r fairshare_redist, echo = FALSE, warning = FALSE}

local_redist <- lm_robust(data = df, 
           redistSelf ~ localFairShare +  p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight,
           fixed_effects = region, clusters = PCON19CD)

region_redist <- lm_robust(data = df, 
                           redistSelf ~ regionFairShare + p_edlevel + age + male + 
                             p_socgrade  + p_gross_household + leftRight,
                           fixed_effects = region, clusters = PCON19CD)

london_local_redist <- lm_robust(data = df, 
           redistSelf ~ london_local_fair_share  + p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight, 
           fixed_effects = region, clusters = PCON19CD)

london_region_redist <- lm_robust(data = df,
                                  redistSelf ~ london_region_fair_share + p_edlevel + age + 
                                    male + p_socgrade + p_gross_household + leftRight,
                                  fixed_effects = PCON19CD, clusters = PCON19CD)
modelsummary(list(local_redist, region_redist, 
                  london_local_redist, london_region_redist),
             stars = TRUE,
             output = "markdown",
             coef_map = c("localFairShare" = "Local area gets fair share", 
                             "regionFairShare" = "Region gets fair share",
                             "london_local_fair_share" = "London-local fair share",
                              "london_region_fair_share" = "London-region fair share",
                             "p_edlevel" = "Education",
                             "age" = "Age",
                             "male" = "Male",
                             "p_socgrade" = "Social grade",
                             "p_gross_household" = "Household income",
                             "leftRight" = "Left-right self-placement"),
             coef_omit = "factor",
             title = "Perceptions of geographic redistribution and preferences over redistribution",
             notes = "Linear models with region fixed effects.
            Dependent variable is preferences over redistribution on 1-10 scale.")

```


This set of models looks at the relationship between redistribution preferences and three other variables that measure perceptions of overall inequality, of geographic deservingness, and of perceptions of an area's relative standing in the UK's overall wealth distribution. I intend these models will help us understand whether the strong relationship, found above, between redistribution preferences and perceptions for deservingness are independent of - and survive when controlling for - concerns about inequality more generally and geographic inequality in particular. 

The two new variables in this section are the overall-inequality and geographic-inequality variables. The former is an item that asks respondents about whether they believe "economic inequality" is changing in the UK; responses ranged on a five-point scale, from "getting much lower" to "getting much higher." The majority of respondents indicated they believe inequality is worsening in the UK. The variable on \textit{geographic} inequality derives from an item that asks respondents to rank their own local community in the UK's overall wealth distribution. Responses cold range from 1 to 100, with 1 indicating the community is among the poorest, while 100 indicates the community is among the richest and 50 indicates the community is at the median. Just over half of respondents rank their areas as below the median in wealth; the distribution of the variable is roughly normal. 


```{r area_rich_poor}

redist_changeInequality <- lm_robust(data = df, 
           redistSelf ~ changeInequality + 
             p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight,
           fixed_effects = PCON19CD, clusters = PCON19CD)

redist_changeInequality_areaRichPoor <- lm_robust(data = df, 
           redistSelf ~ changeInequality + areaRichPoor_scale + 
             p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight,
           fixed_effects = PCON19CD, clusters = PCON19CD)

redist_changeInequality_areaRichPoor_interact <- lm_robust(data = df, 
           redistSelf ~ changeInequality*scale(areaRichPoor) + 
             p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight,
           fixed_effects = PCON19CD, clusters = PCON19CD)


redist_localFairShare_areaRichPoor_interact <- lm_robust(data = df, 
           redistSelf ~ localFairShare*areaRichPoor_binary_rich + changeInequality + 
             p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight,
           fixed_effects = PCON19CD, clusters = PCON19CD)

modelsummary(redist_changeInequality_areaRichPoor, 
             stars = TRUE,
             output = "markdown")

redist_changeInequality_areaRichPoor_localFairShare <- lm_robust(data = df, 
           redistSelf ~ changeInequality + scale(areaRichPoor) + localFairShare + 
             p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight,
           fixed_effects = PCON19CD, clusters = PCON19CD)




modelsummary(list(redist_changeInequality,
                  redist_changeInequality_areaRichPoor, 
                  redist_changeInequality_areaRichPoor_interact,
                  redist_changeInequality_areaRichPoor_localFairShare), 
             stars = TRUE,
             output = "markdown",
             coef_map = c(
               "changeInequality" = "Perceived change in inequality",
               "scale(areaRichPoor)" = "Local area is relatively poor/rich", 
               "changeInequality:scale(areaRichPoor)" = "Change in inequality * local poor/rich",
               "localFairShare" = "Local area gets fair share", 
               "leftRight" = "Left-right placement"
             ))


coef_map = c(
               "scale(areaRichPoor)" = "Relative perceptions of local economy as rich-poor",
               "localFairShare" = "Local area gets fair share",
               "p_edlevel" = "Education",
               "age" = "Age",
               "male" = "Male",
                "p_socgrade" = "Social grade (higher values = lower)",
               "p_gross_household" = "Income", 
               "leftRight" = "Left-right placement"
             ))


redist_changeInequality_areaRichPoor_lm <- lm(data = df, 
           redistSelf ~ localFairShare*areaRichPoor_binary_rich + 
             p_edlevel + age + male + p_socgrade + 
             p_gross_household + leftRight + as.factor(region))

plot((ggpredict(redist_changeInequality_areaRichPoor_lm,
                terms = c("localFairShare", "areaRichPoor_binary_rich"))))


```



Let us first look at models 1 and 2. In each model, the key coefficients are negative and significant: As respondents believe their local area or region receive increasingly "fair" shares of central government spending, their preferences for generalized redistribution decrease. Models 3 and 4 look at respondent's perceptions that London receives an undue amount of central spending - this may capture perceptions of peripheralness, or perhaps place-based economic resentment. In both models, the coefficients are positive and significant: Respondents who believe London receives too much spending are also more favorable towards redistribution. Across all four models, the coefficients in the local-area models (1 and 3) are larger than in the regional models (2 and 4); this suggests that local areas may be \textit{more} salient to voters than are regions. The broad takeaway from these models, as I see it, is that perceptions of deservingness do correspond with preferences over redistribution. 

I now look at the relationship between perceptions of economic inequality and preferences over redistribution. The BES included a series of questions that asked respondents about their perceptions of the annual economic performance of their own local area, their region, and London. Replicating the structure above, models 1 and 2 use perceptions of local and regional economic performance as their key independent variable, while models 3 and 4 use the difference between perceptions of London's economy and perceptions of the local and regional economies respectively. All models use redistribution preferences as the dependent variable. These latter models therefore inquire directly into the relationship between perceptions of geographic-economy inequality and redistribution. I again use a suite of demographic covariates and regional fixed effects. 
```{r echo = FALSE, warning = FALSE}
local_econ_mod <- (lm_robust(data = df, 
                   redistSelf ~ localEcon + p_edlevel + age + 
                     male + p_socgrade + p_gross_household + leftRight, 
                   fixed_effects = region))
region_econ_mod <- (lm_robust(data = df, 
                   redistSelf ~ regionEcon + p_edlevel + age + 
                     male + p_socgrade + p_gross_household + leftRight, 
                   fixed_effects = region))
london_local_econ_mod <- (lm_robust(data = df, 
                   redistSelf ~ london_local_econ + p_edlevel + age + 
                     male + p_socgrade + p_gross_household + leftRight, 
                   fixed_effects = region))
london_region_econ_mod <- (lm_robust(data = df, 
                   redistSelf ~ london_region_econ + p_edlevel + age + 
                     male + p_socgrade + p_gross_household + leftRight, 
                   fixed_effects = region))
modelsummary(list(local_econ_mod, region_econ_mod, 
                  london_local_econ_mod, london_region_econ_mod),
             stars = TRUE,
             output = "latex",
             coef_map = c("localEcon" = "Local econ perception", 
                             "regionEcon" = "Regional econ perception",
                             "london_local_econ" = "London-local econ perception",
                              "london_region_econ" = "London-region econ perception",
                             "p_edlevel" = "Education",
                             "age" = "Age",
                             "male" = "Male",
                             "p_socgrade" = "Social grade",
                             "p_gross_household" = "Household income",
                             "leftRight" = "Left-right self-placement"),
             coef_omit = "factor",
             title = "Perceptions of geographic inequality and preferences over redistribution",
             notes = "Linear models with region fixed effects.
            Dependent variable is preferences over redistribution on 1-10 scale.")

```

The results here are similar to those in the models in table 1. In models 1 and 2, positive perceptions of the local and regional economy correlate negatively with preferences for higher redistribution. The coefficient is larger, and more strongly significant, for the local economy than for the regional one. Models 3 and 4 show that perceptions that London's economy is outpacing the local or regional one associate strongly with preferences for \textit{higher} redistribution. I take this as evidence that perceptions of geographic inequality - which may map onto a deeper concept of "perceived peripheralness," or something along those lines - play into redistribution preferences. This seems a fascinating relationship to explore further, as I briefly develop below. 

Finally, table 3 looks at the relationship between perceptions of the economy and perceptions of geographic redistribution. Models 1 and 2 use the local and regional fair-share measures respectively as the dependent variable, and perceptions of local and regional economic performance as the independent variables. The positive and significant coefficients indicate that, as respondents believe their own local or regional economies are doing better, they perceive those areas as receiving their fair share of spending from the central goverment. (This may suggest that individuals believe poorer places \textit{should} get a higher share of funding). 

Models 3 and 4 change the dependent variables to, respectively, the difference between the perceived share of funding going to London and going to one's local area or region. Higher values indicate the respondent believes London is receiving an undue share, relative to their own local area or region. In both cases, the coefficients are negative and significant: Better perceptions of the local/regional economy correspond negatively with the belief that London receives an undue share of revenue. This suggests, in turn, that place-based economic resentment is at least in part a product of individuals' perceptions of local/regional economic performance. 

```{r echo = FALSE, warning = FALSE}

# look at the relationship between perceptions of the economy and perceptions of geographic unfairness

local_econ_fairshare <- lm_robust(data = df, 
                                  localFairShare ~ localEcon + p_edlevel + age + 
                                    male + p_socgrade + p_gross_household + leftRight, 
                                  fixed_effects  = region)


region_econ_fairshare <- lm_robust(data = df, 
                                  regionFairShare ~ regionEcon + p_edlevel + age + 
                                    male + p_socgrade + p_gross_household + leftRight, 
                                  fixed_effects  = region)

local_econ_london_local_fairshare <- lm_robust(data = df, 
                                  london_local_fair_share ~ localEcon + p_edlevel + age + 
                                    male + p_socgrade + p_gross_household + leftRight, 
                                  fixed_effects  = region)

region_econ_london_region_fairshare <- lm_robust(data = df, 
                                  london_region_fair_share ~ regionEcon + p_edlevel + age + 
                                    male + p_socgrade + p_gross_household + leftRight, 
                                  fixed_effects  = region)

modelsummary(list(local_econ_fairshare, region_econ_fairshare,
                  local_econ_london_local_fairshare, 
                  region_econ_london_region_fairshare),
             stars = TRUE,
             output = "markdown",
             coef_map = c("localEcon" = "Local economic perception", 
                             "regionEcon" = "Regional economic perception",
                             "p_edlevel" = "Education",
                             "age" = "Age",
                             "male" = "Male",
                             "p_socgrade" = "Social grade",
                             "p_gross_household" = "Household income",
                             "leftRight" = "Left-right self-placement"),
            title = "Correlation between perceptions of economic performance and geographic redistribution")


```

\section{Next steps}
\begin{enumerate}
\item Look at the relationship between the fair-share variables and objective conditions. Do respondents correctly perceive that their areas receive less than they should (however that may be defined) from the central government? This will depend on the availability of official statistics on distributional formulas. 

\item Look at movers. Do people who move to a poorer place become more favorable to redistribution? This is empirically feasible due to the panel structure of the data, plus the geographic identifiers. 

\item Investigate the relationship with both economic perceptions and objective economic conditions. When the local economy begins to suffer, do individuals come to perceive their areas as receiving less than their fair shares of funding? 

\item Look at the relationship with vote choice. Previous work (Green and Fieldhouse 2021) has suggested that, in the British context, those who feel senses of place-based economic resentment were more likely to vote for Brexit. Does this apply to general elections as well? Was place-based economic resentment, for instance, a driver of Labour-to-Conservative switching in the 2019 general election? 

\item Look at geographic correlates of feelings of economic resentment. Does distance from London play a role? Distance from the closest major city? 

\item Look at the relationship between the economic composition of the place - in particular the extent to which the place is "digitalized," or otherwise participates in the knowledge economy - and perceptions of geographic redistribution. Are places whose economic composition is quite "left-behind" more resentful towards London's share of centralized spending? This expands, in a sense, the scope of investigation into economic compositions. The research question could be something like: How does peripherality/left-behindness influence views over geographic redistribution? 

\end{enumerate}